---
title: "Homework 2: Health Economics"
author: "Nikhita Gandhe"
date: "02/18/2025"
format: pdf
---

# Please Find my GitHub Repository Link Here: 
https://github.com/nsgand2/ECON-470-HW2.git  

```{r setup, include=FALSE}
# loading libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, data.table, gdata)
library(readr)     # For reading data
library(knitr)
library(tidyr)
library(MatchIt)    
library(WeightIt)   
library(sandwich)   
library(lmtest)     
library(Matching)   
  
load("C:/Users/Nikhita Gandhe/Documents/GitHub/ECON-470-HW2/submission3/hw_workspace.Rdata")
```

# Question 1: Number of Hospitals Filing Multiple Reports
How many hospitals filed more than one report in the same year? Show your answer as a line graph of the number of hospitals over time.
```{r}
#| echo: FALSE
  
  # graphing
  ggplot(data = total_report_count, aes(x=year, y=total)) +
    geom_line()
  
```

# Question 2: Unique Hospital IDs
After removing/combining multiple reports, how many unique hospital IDs (Medicare provider numbers) exist in the data?

```{r}
#| echo: false

tibble(unique_hospital_ids = nrow(unique_provider_nos))

```


# Question 3: Distribution of Total Charges by Year
What is the distribution of total charges in each year? Show your results with a violin plot.

```{r}
#| echo: FALSE

print(violin_plot)

```


# Question 4: Distribution of Estimated Prices by Year
What is the distribution of estimated prices in each year? Show your results with a violin plot.

```{r}
#| echo: FALSE
ggplot(hcris_filtered, aes(x = year, y = price)) +
  geom_violin(fill = "red", alpha = 0.5, adjust = 1.5) +  # Smooth density estimation
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.3) +  # Boxplot for additional clarity
  scale_y_continuous(trans = "log10") +  # Log scale to handle skewed data
  theme_minimal() +
  labs(title = "Distribution of Estimated Prices by Year",
       x = "Year",
       y = "Estimated Price (log scale)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate labels for readability

```

# Question 5: Average Price for Penalized vs. Non-Penalized Hospitals
Calculate the average price among penalized versus non-penalized hospitals.

```{r}
#| echo: FALSE

print("Average Price by Penalty Status (2012):")
print(average_prices)

```


# Question 6 Split hospitals into quartiles based on bed size.
Provide a table of the average price among treated/control groups for each quartile.
```{r}
#| echo: FALSE

kable(average_prices_by_quartile, caption = "Average Price by Penalty Status Across Bed Size Quartiles")


```

# Question 7 Find the average treatment effect using each of the following estimators, and present your results in a single table:

```{r}
#| echo: FALSE
kable(results, caption = "Estimated Treatment Effects by Method")

```

# Question 8 With these different treatment effect estimators, are the results similar, identical, very different?
The results are very different between the different treatment effect estimators.

# Question 9 Do you think you’ve estimated a causal effect of the penalty? Why or why not? (just a couple of sentences)
Due to a high variance in between the estimates, I was unable to estimate a causal effect for the penalty.

# Question 10 Briefly describe your experience working with these data (just a few sentences). Tell me one thing you learned and one thing that really aggravated or surprised you.
I learned that I can easily work on a r file in VS Code and then move everything to a qmd file if I just load it using this line of code which I run in the r file and in the beginning of the qmd file, load("C:/Users/Nikhita Gandhe/Documents/GitHub/ECON-470-HW2/submission3/hw_workspace.Rdata"). One thing that really aggravated me was that despite using the same code as other classmates, the estimates we obtained had a very high variance. 
